\textbf{International Transport costs:\\New Findings from modeling additive costs}\\Online Appendix (Not for Publication)

### 2021-07-09 redaction/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
Modids sur les parties liées aux IV
rangements et modif mineures (OA)

Trade costs are central to international economic analysis. In particular, they are
known to be a major obstacle to international economic integration and international trade
flows. Defined as the costs associated with the exchange of goods across
national borders, trade costs are usually split between transaction costs, policy costs, time costs and transport costs \textit{per se}. Many argue that the latter have substantially
decreased with technological advance in transportation, infrastructure development
and new communication technologies. However, several empirical evidence
suggest that transport costs remain large and deserve attention.\footnote{\cite{anderson_wincoop_jel} find that around 30\% of international trade costs are attributable to transport costs. Equivalently, international transportation costs represent a 21\% markup over production costs. Further, the sizable elasticity of trade with respect to freight costs obtained by \cite{Behar_Venables} (around -3) testifies to the impact of transport costs on trade flows. \cite{Disdier_Head08}, among others, point out that distance as an incompressible cost is still a very important determinant of trade flows.} In addition, progress in transportation techniques (which reduces the constant-dollar cost of transport) should be compared with technical progress in production (which reduces the constant-dollar cost of production) to get a clear picture of the evolution
of the transport costs burden. The latter also depends on whether the cost is in proportion of the value of goods traded (hence does not distort relative prices) or a cost per unit traded (the additive form conversely affecting relative prices). If a complex task, having a clear view of the international transport cost burden is of key importance, since the answers to several important (both academic and policy) economic questions depend in large part on how much distance-based costs affects trade, such as, among others, the dynamics
of agglomeration, the profitability of outsourcing or the gains from regional liberalization. The main contribution of the paper is to bring new insights on this issue.\smallskip

Our paper contributes to the literature by providing a thorough empirical investigation of international transport costs over time based on US imports flows from 1974 to 2019. Importantly, our empirical setting carefully distinguishes between multiplicative and additive transport costs, which proves to be key in the characterization of international trade patterns. While most of the international trade literature has modeled trade costs as an ad-valorem tax equivalent (i.e. as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis in Samuelson's (1954) terminology\cite{samuelson1954}),\footnote{Strictly speaking, the ``iceberg cost'' includes both the ad-valorem dimension of trade cost and the fact that these costs are paid in terms of the good that is traded.
As this last element is irrelevant in our case, we use the terms ``iceberg'' and ``ad-valorem'' interchangeably, as is commonly done in the literature.} a recent strand of the empirical literature points to the existence of an additive component, that is, a cost per unit traded (see, among others, \cite{Irrazabal_2015}, or \cite{martin2012}). Based on US sectoral data, our own estimates also point out the importance of additive costs in international transport costs. Contributing one step further, we provide estimates of the size of both additive and ad-valorem transports costs by sector (at the 3-digit level) and country of origin on a yearly basis over 1989-2019 and by transport mode (air or maritime transport).\smallskip

>>>>

Trade costs are central to international economic analysis. In particular, they are known to be a major obstacle to international economic integration and international trade flows. Defined as the costs associated with the exchange of goods across
national borders, trade costs are usually split between transaction costs, policy costs, time costs and transport costs \textit{per se}. Many argue that the latter have substantially decreased with technological advance in transportation, infrastructure development
and new communication technologies. However, several empirical evidence
suggest that transport costs remain large and deserve attention.\footnote{\cite{anderson_wincoop_jel} find that around 30\% of international trade costs are attributable to transport costs. Equivalently, international transportation costs represent a 21\% markup over production costs. Further, the sizable elasticity of trade with respect to freight costs obtained by \cite{Behar_Venables} (around -3) testifies to the impact of transport costs on trade flows. \cite{Disdier_Head08}, among others, point out that distance as an incompressible cost is still a very important determinant of trade flows.} In addition, progress in transportation techniques (which reduces the constant-dollar cost of transport) should be compared with technical progress in production (which reduces the constant-dollar cost of production) to get a clear picture of the evolution
of the transport costs burden. The latter also depends on whether the cost is in proportion of the value of goods traded (hence does not distort relative prices) or a cost per unit traded (the additive form conversely affecting relative prices). If a complex task, having a clear view of the international transport cost burden is of key importance, since the answers to several important (both academic and policy) economic questions depend in large part on how much distance-based costs affects trade, such as, among others, the dynamics
of agglomeration, the profitability of outsourcing or the gains from trade liberalization. The main contribution of the paper is to bring new insights on this issue.\smallskip

Our paper contributes to the literature by providing a thorough empirical investigation of international transport costs over time based on US imports flows from 1974 to 2019. Importantly, our empirical setting carefully distinguishes between multiplicative and additive transport costs, which proves to be key in the characterization of international trade patterns. While most of the international trade literature has modeled trade costs as an ad-valorem tax equivalent (i.e. as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis in Samuelson's (1954) terminology\cite{samuelson1954}),\footnote{Strictly speaking, the ``iceberg cost'' includes both the ad-valorem dimension of trade cost and the fact that these costs are paid in terms of the good that is traded.
As this last element is irrelevant in our case, we use the terms ``iceberg'' and ``ad-valorem'' interchangeably, as is commonly done in the literature.} a recent strand of the empirical literature points to the existence of an additive component, that is, a cost per unit traded (see, among others, \cite{Irrazabal_2015}, or \cite{martin2012}). Based on US sectoral data, our own estimates also point out the importance of additive costs in international transport costs. Contributing one step further, we provide estimates of the size of both additive and ad-valorem transports costs by sector (at the 3-digit level) and country of origin on a yearly basis over 1974-2019 and by transport mode (air or maritime transport).\smallskip
